---
title: "Confounding and effect modification"
author: "Fayssal Ali"
date: "2022-10-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libraries}

#Assignment 1
library(tibble)
library(dplyr)
library(epiR)
library(readr)
library(DescTools)
library(epitools)
library(ggplot2)
library(ggstatsplot)
library(crosstable)


```
#Assignment 2:

Question 1 Risk Difference:
 (π11/π1+) - (π21/π2+)

Question 2 Relative Risk:
 (π11/π1+) / (π21/π2+)

Question 3 Odds Ratio:
(π11/π12) * (π22/π21)




#Assignment 3    
```{r}

#X^2, df=9
qchisq(.95, df=9)
#Critical Value = 16.9
#N(0,1)
#H0: u<= u0
#H1: u > u0
qnorm(.975,lower.tail = T)
#Critical Value=1.96
#F3,7
qf(.95,3,7)
#Critical Value= 4.35

```

Relative Risk would not be an appropriate measure of association for this study. 
Case-Control studies have a fixed disease and non-diseased population with random  
exposure and non-exposure. Therefore, estiamtes can only be made on P(E|D) and P(E|non-D)
This probabilty is also the estimated OR. 


The odds of mortality in patients with severe pancreas/duodenal injury undergoing exploratory laparotomy and trauma pancreaticoduodenectomy are 38% less than those with exploratory laparotomy alone. However, the CI interval includes 1 and the p-value for this statistic is <0.05, therefore this OR is not significant.

The odds of major complications in patients with severe pancreas/duodenal injury undergoing exploratory laparotomy and trauma pancreaticoduodenectomy are 244% greater than those with lapartomy alone. The OR CI does not include 1 and the p-value <0.05 so this result is statistically meaningfull.

The odds of venous thromboembolism in patients with severe pancreas/duodenal injury undergoing exploratory laparotomy and trauma pancreaticoduodenectomy are 386% greater than those with lapartomy alone. The OR CI does not include 1 and the p-value <0.05 so this result is statistically meaningfull.


```{r}

#Summary statistics 
xtabs( ~ mcollege, data = School)
xtabs( ~ income, data = School)
School$mcollege <- factor(School$mcollege, levels = c("no", "yes"))
School$income <- factor(School$income, levels = c("low", "high"))

#Table Summary:
crosstable::crosstable(
  School,
  mcollege,
  by=income,
  total = "both",
  percent_pattern="{n}"
) %>%
  crosstable::as_flextable(
    keep_id = FALSE
  )

#Graphical Visulization:
ggbarstats(School,
           mcollege,
           income,
           type = "parametric",
           xlab = "income",
           title = "income by college completion"
)

```

```{r}

new <- table(School$mcollege,School$income)
new
epiR::epi.2by2(new)
```
#OR= 4.06 (3.2,4.82)

```{r}

chisq.test(new)

```

#X^2= 288.44; p-value<0.001
#Reject the null hypothesis and conclude an aossociation does exsist betnween mothers college and income

```{r}

xtabs(~ ethnicity, data=School)
#Create stratfied table
strat <- table(School$mcollege, School$income, School$ethnicity)

#Conduct Woolf test of Homogeneity on the stratified tables
DescTools::WoolfTest(strat)

```
#pvalue= 0.89 and x2 = 0.225 fail to reject hypothesis of homogenous association between the stratified groups
#Not homogeneous association, Check if Ethnicity is potential confounder

```{r}
#Check for confounding:
#create table with X and Z
eth <- table(School$mcollege, School$ethnicity)
eth
#test for association 
chisq.test(eth) #61.84 and p <0.001
#Reject the null and conclude there is an association between mother's collge and ethnicity
new
#Ethnicity is not on the casual pathway of mother's college

#Conduct measuers of association testing ofr each level of ethnicity:

#afam
af <- matrix(
  c(582,55,107,42),
  nrow = 2,
  dimnames = list(
    c("no", "yes"),
    c("low", "high")
  )
)

epiR::epi.2by2(af)
#Hispanic
his <- matrix(
  c(718,33,129,23),
  nrow = 2,
  dimnames = list(
    c("no", "yes"),
    c("low", "high")
  )
)

epiR::epi.2by2(his)

#Other
ot <- matrix(
  c(1792,194,760,304),
  nrow = 2,
  dimnames = list(
    c("no", "yes"),
    c("low", "high")
  )
)
epiR::epi.2by2(ot)
#The OR for each level of ethnicity seem to be very similar

#Conduct Mantel-Haenszel test
epiR::epi.2by2(strat)

#MH OR= 3.77 (3.16,4.48)
```

#Check the rule of thumb:
#(4.06-3.77)/4.06 =0.07 <0.1
#Report the crude OR 

```{r}
xtabs(~private, data=Aus)
xtabs(~age, data=Aus)
xtabs(~gender, data=Aus)
xtabs(~lchronic, data=Aus)

Aus$private <- factor(Aus$private, levels = c("no","yes"))

Aus$lchronic <- factor(Aus$lchronic, levels = c("no","yes"))

Aus$gender <- factor(Aus$gender, levels = c("female","male"))

#Table Summary:
crosstable::crosstable(
  Aus,
  private,
  by=gender,
  total = "both",
  percent_pattern="{n}"
) %>%
  crosstable::as_flextable(
    keep_id = FALSE
  )

crosstable::crosstable(
  Aus,
  private,
  by=lchronic,
  total = "both",
  percent_pattern="{n}"
) %>%
  crosstable::as_flextable(
    keep_id = FALSE
  )

#Graphical Visulization:
ggbarstats(Aus,
           private,
           gender,
           type = "parametric",
           xlab = "gender",
           title = "private insurance by gender distribution"
)

ggbarstats(Aus,
           private,
           lchronic,
           type = "parametric",
           xlab = "lchronic",
           title = "private insurance by chronic condition"
)

```

```{r}

#Age
pr <- glm(private~age, data=Aus, 
          family = binomial(link = "logit"))
summary(pr)$coeff

#Gender
ge <-  glm(private~gender, data=Aus, 
           family = binomial(link = "logit"))
summary(ge)$coeff

#Lchronic
lc <- glm(private~lchronic, data=Aus, 
          family = binomial(link = "logit"))
summary(lc)$coeff
```

```{r}

full<- glm(private~age+gender+lchronic, data=Aus, 
          family = binomial(link = "logit"))
summary(full)$coeff
#Model
# Y= 0.24 - 0.74x1 - 0.32x2 - 0.24x3
#Intercept: 0.25 = the log odds of private insurance for those who are female, no chronic condition, and age=0
#B1: -0.74= the log odds of private insurance for a one-year increase of age, adjusted for gender and prescence of chronic condition 
#B2: -0.32= the log odds of private insurance comparing females and males adjusted for age andprscence of chronic condition 
#B3: -0.24= the log odds of private insurance comparing those with a chronic condition and those without adjusted for age and gender
```

```{r}

#probability of insurance given---- age= 30(30/100),female(0) no chronic(0)
exp(0.25 - (.74*.32))/(1+ exp(0.25 - (.74*.32)))

#50.33% of having private insurance given a female age 30 with no chronic condition 
```
```{r}
#Question 7:
#Check for effect modification:

full2 <- glm(private~gender+lchronic+gender*lchronic, data=Aus, 
         family = binomial(link = "logit"))
summary(full2)$coeff

#Zstat of 0.9 and pvalue of 0.37 
#Fail to reject the null hypothesis and conclude lchronic does not modify the association 
#between gender and private insurance

#Check for confounding 
w <- glm(private~gender, data=Aus, 
         family = binomial(link = "logit"))
summary(w)$coeff
exp(coefficients(w))
exp(confint.default(w, level = 0.95))
#OR= 0.79 (0.71,0.89) comaring females to males in the crude model

adj <- glm(private~gender+lchronic, data=Aus, 
           family = binomial(link = "logit"))

summary(adj)$coeff
exp(coefficients(adj))
exp(confint.default(adj, level = 0.95))
#OR= 0.79 (0.71,0.88) comparing females to males when adjusted for presence of a chronic condition 

#Since OR and CI are very similar, can report the crude numbers

#Can also confirm/check with table methods:
p <- table(Aus$gender, Aus$private, Aus$lchronic)
p
epiR::epi.2by2(p)

```
















