---
title: "Logistic Models  with continuous variables" 
author: "Fayssal"
date: "2022-10-12"
output: html_document
---
#Investigating the association between gang membershup and HIV sero-converesion, while 
#age of onset smoking will be considered as an effect modifier in South Africa and Brazil


#Part 1(South Africa)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
africa <- gangs_hiv_africa
```



#Factor appropriately

```{r}
africa <- gangs_hiv_africa

africa$Gang <- factor(
  africa$Gang, levels = c("No", "Yes")
)


africa$HIV <- factor(
  africa$HIV, levels = c("Negative", "Positive")
)

```




```{r }
model3 <- glm(
  formula= HIV ~ Gang + Age + Gang * Age,
  data = africa,
  family = binomial(link = "logit")
)
summary(model3)$coefficients
```

#log= 0.918-0.724x+0.005z-0.059xz
#Zstat for B3(interaction)=-.49 and pvalue=0.62
#Fail to reject the null and conclude onset smoking age DOES NOT modifify the association between gang membership  and HIV status



```{r}
quantile(
 africa$Age,
  probs = c(0.25, 0.5, 0.75)
)
```
#Graphical Visualization 
```{r}
emmeans::emmip(
  model3,
  ~ Gang | Age,
  at = list(Age = c(12, 14, 16))
)
```
#Part2
```{r}
brazil <- gangs_hiv_brazil

brazil$Gang <- factor(
  brazil$Gang, levels = c("No", "Yes")
)


brazil$HIV <- factor(
  brazil$HIV, levels = c("Negative", "Positive")
)
```

```{r}
model4 <- glm(
  formula= HIV ~ Gang + Age + Gang * Age,
  data = brazil,
  family = binomial(link = "logit")
)
summary(model4)$coefficients
```
#log= 10.13+29.62x-0.43z-2.55xz
#Zstat for B3(interaction)=-1.98 and pvalue=0.047
#Reject the null and conclude onset smoking age modififies the association between gang membership  and HIV status

```{r}
quantile(
brazil$Age,
  probs = c(0.25, 0.5, 0.75)
)
```
```{r}
emmeans::emmip(
  model4,
  ~ Gang | Age,
  at = list(Age = c(13, 14, 22))
)
```

#Model for each value
```{r}
model4.1  <- emmeans::emmeans(
    model4,
    ~ Gang | Age,
    at = list(Age = c(13, 14, 22))
  )
```
#Find the Estimate for each value (13,14,22)
```{r}
emmeans::contrast(
  model4.1,
  "consec",
  name = "Gang",
  infer = c(TRUE, TRUE),
  level = 0.95
)
```
#Exponintatie each age
```{r}
#13
exp(c(-3.59, -6.67, -0.512))
#14
exp(c(-6.14, -8.99, -3.295))
#22
exp(c(-26.58, -47.91, -5.248))

```


#As age increases so does the OR decrease!







