---
title: "Novel Antibotics study with covariates"
author: "Fayssal"
date: "2022-10-12"
output: html_document
---

Investigating the effect of a novel antibiotic compared to an established antibiotic on the development of surgical surgery, while also considering hollow-viscus injury as a potential effect modifier. Hollow Viscus injury will have three categorical levels 
(None, SBI, LBI)



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Upload the Dataset
emp <- read_csv("empirical_injury_infection_data.csv")


```

#Factor accordingly 

```{r emp}
emp$Injury <- factor(
  emp$Injury,
  levels = c("None", "SBI", "LBI")
)

emp$Infection <- factor(
  emp$Infection,
  levels = c("No", "Yes")
)

emp$Group <- factor(
  emp$Group,
  levels = c("Established", "Novel")
)
```

## Build the Model

#Refrence Group= No Injury 

```{r }
model2 <- glm(Infection~ Group + Injury + Group*Injury,
              data=emp,
              family = binomial(link = "logit")
)
summary(model2)$coefficient
```
Model:
log=-2.33 0.28x 0.20z1 +0.16z2 -0.54xz1 -1.76xz2
log= -2.33 +0.20z1 +0.16z2 +(0.28z1 -0.54z2 -1.76)x
#Test the reduced model without the interaction 

```{r}
model2.1 <- glm(Infection~ Group + Injury,
              data=emp,
              family = binomial(link = "logit")
)
summary(model2.1)$coefficient
```
#LRT test
#Testing H0: B4=B5=0 (states Z is not an effect modifier)


```{r}

lmtest::lrtest(model2.1, model2)
```
#2 degrees of freedom(6-4)
#G^2 stat= 38.329 and pvalue= <0.001
#Reject the null hypothesis and conclude Injury does modify the association between anitboitic group and surgical infection. 

#Log odds ratio for all 3 levels:


#Show the individual results of hypothesis testing and CI
```{r}
em_grid2 <- emmeans::emmeans(
  model2,
  ~ Group | Injury
)

emmeans::contrast(
  em_grid2,
  "consec",
  name = "Injury",
  infer = c(TRUE, TRUE),
  level = 0.95
)
```
#Antilog calculations
```{r}
exp(c(0.283, -0.0557, 0.6226))
exp(c(-0.253, -0.6014, 0.0964))
exp(c(-1.477, -1.9390, -1.0144))
```
#Conclusion: For those with no injury or SBI, there is not enough evidence to show an association between antibiotic group and surgical infection



#Graphical representation 
```{r}
emmeans::emmip(
  model2,
  ~ Group | Injury
)
```

