---
title: "Recurrence of breast-cancer"
author: "Fayssal"
date: "2022-10-19"
output: html_document
---


Investigating the effect of the presence of nodecaps on the recurrence of breast-cancer from the UCI data repository, while also considering tumor malignancy grade as a possible effect modifier. 
Malignancy grade will have three categorical levels (1,2,3). 

Background:

Breast cancer is the second leading cause of death among women worldwide. On Average 7-11% of women with early breast cancer will have a recurrence. Specifically the same cancer has reappeared, this is not a new form of cancer. One of the major signs of recurrence is the size of the malignant tumor. 

Most malignant tumors are graded as  1, 2, 3, or 4. In grade 1 tumors, the cells look close to normal. The higher the number, the more abnormal the cells look. Grade 3-4 tumors look most abnormal.
(Our dataset has tumor grades from 1-3)
Finally, the Node-caps in our data-set distinguish those that cancer has spread to the lymph nodes.

We will be exploring dataset found in the UCI repository there are 286 observations and with 6 variables. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(epitools)
library(DescTools)
library(epiR)
library(emmeans)
library(lmtest)


breast_cancer <- read_csv("breast-cancer.csv", 
     col_names = FALSE)

bc <- breast_cancer[c(1,2,3,6,7,8)]

colnames(bc) <- c("Cancer","AgeGroup","Menopause","nodecaps","malignancy","radiation")

bc <- bc[!(bc$nodecaps=="?"),]
```



```{r cars}
bc$Cancer <- factor(
  bc$Cancer, levels =c("no-recurrence-events","recurrence-events")
)

bc$malignancy <- factor(
  bc$malignancy, levels =c("1","2","3")
)


bc$nodecaps <- factor(
  bc$nodecaps, levels =c("no","yes")
)

bc$AgeGroup <- factor(
  bc$AgeGroup, levels =c("20-29","30-39","40-49","50-69", "60-69","70-79")
)

bc$Menopause <- factor(
  bc$Menopause, levels =c("premeno", "ge40", "lt40")
)


bc$radiation <- factor(
  bc$radiation, levels =c("left", "right")
)

```


```{r }
#Summary of raw numbers
xtabs( ~ AgeGroup,data = bc)
xtabs( ~ malignancy , data = bc)
xtabs( ~ Cancer, data = bc)
xtabs(~nodecaps, data= bc)
xtabs(~radiation, data=bc)
xtabs(~Menopause, data=bc)
```

```{r}
#Create pooled table
can <- table(bc$Cancer, bc$nodecaps)
can
```
2X2
```{r}
crosstable::crosstable(
  bc,
  Cancer,
  by=nodecaps,
  total = "both",
  percent_pattern="{n}"
) %>%
  crosstable::as_flextable(
    keep_id = FALSE
  )
```
```{r}
epiR::epi.2by2(can)
```
#The odds of no-reccurrence of breast cancer in the no node-caps group is 316% higher 
#when comapred to the recurrence-events


#The odds of reccurrece of breast cancer in the yes node-caps group is 316% higher when comparred to no-reccurrence

```{r}
mal <- table(bc$Cancer, bc$malignancy)
mal

crosstable::crosstable(
  bc,
  Cancer,
  by=malignancy,
  total = "both",
  percent_pattern="{n}"
) %>%
  crosstable::as_flextable(
    keep_id = FALSE
  )
```

```{r}
#Stratified Table
strat <- table(bc$Cancer, bc$nodecaps, bc$malignancy)
strat


DescTools::WoolfTest(strat)



```

#MH X2= 1.64 with a p-value of 0.44
#Fail to reject the null hypothesis and conclude homogonous association between the straitifed tables
#Not an effect modifier 

#Check for confounding 

```{r}
mal <- table(bc$nodecaps, bc$malignancy)
mal

chisq.test(mal)
```

#Reject the null and conclude there is an association between node-caps and malignancy grade
#Check if malignancy grade is the causal pathway of node-caps
#Many articles suggest node-caps found in the lymph  nodes are a result of the malignant tumor spreading

#Malignancy is not a effect modifier or confounder
#Report crude numbers 

```{r}
ggbarstats(bc,
          Cancer,
           nodecaps,
           type = "parametric",
           xlab = "nodecaps",
           title = "nodecaps by carcinoma diagnosis"
)

```

```{r}
ggbarstats(bc,
          nodecaps,
           Cancer,
           type = "parametric",
           xlab = "Carcinoma",
           title = ""
)

```
```{r}
ggbarstats(bc,
          malignancy,
           Cancer,
           type = "parametric",
           xlab = "malignancy",
           title = ""
)
```



```{r}
v <- glm(Cancer~ AgeGroup+malignancy+nodecaps+radiation+Menopause, 
         data=bc, family = binomial(link = "logit"))
summary(v)$coeff
```
Will focus on node-caps and malignancy 

```{r}
full <- glm(Cancer~ nodecaps + malignancy + nodecaps*malignancy, data=bc,
         family= binomial(link = "logit" ))
summary(full)$coeff
```
Model:
log=-1.85 +1.53x 0.42z1 +1.50z2 -0.92xz1 
log= -2.33 +0.42z1 +1.50z2 +(-0.92z1 +1.53)x
#Test the reduced model without the interaction 

```{r}
redu <- glm(Cancer~ nodecaps + malignancy , data=bc,
         family= binomial(link = "logit" ))
summary(redu)$coeff
```

#LRT test
#Testing H0: B4=0 (states Z is not an effect modifier)
```{r}
o <- glm(nodecaps~malignancy, data=bc,
         family= binomial(link = "logit" ))
summary(o)$coeff
```
#Not associated

```{r}
lmtest::lrtest(redu, full)
```
Fail to reject the null hypothesis and conclude that malignancy grade does not effect the association between nodecapes and breast cancer recurrence. (Not an effect modifier)

#Report the crude measures 

```{r}
#Hypthesis testing
em <- emmeans::emmeans(
  mo,
  ~ nodecaps | malignancy
)

emmeans::contrast(
  em,
  "consec",
  name = "malignancy",
  infer = c(TRUE, TRUE),
  level = 0.95
)
```
Note: There are not any obervations with nodecapes present and malaignant grade1.

```{r}
#Malignancy2
exp(c(0.612, -.353,1.58))


#Malignancy3
exp(c(1.533, .525,2.54)) 

emmeans::emmip(
  full,
  ~ nodecaps | malignancy
)
```

#Final Conclusions:

#Malignancy is not an effect modifier or confounder 

#Report the crude measures of association between node-caps and reccurrence of breast cancer for this dataset

#The prescence of a Malignancy grade 3 tumor is a strong predictor of the reccurrence of breast cancer 
#in this dataset

